---
title: "Assignment 1: Creating a Model"
author: "Tom Gibbens-Matsuyama, Haylee Oyler"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

### Goal of this Assignment

Your goal is to implement a simple model of almond yield anomaly response to climate

-   Inputs: daily times series of minimum, maximum daily temperatures and
-   Outputs: maximum, minimum and mean yield anomaly for a input time series

The Lobell et al. 2006 paper will be the source for the transfer function that you will use in your model; specifically look at the equations in table 2.

Here are the expected outputs to check our work:

-   the maximum almond yield anomaly should be approximately 1920 ton/acre
-   the lowest almond yield anomaly should be approximately -0.027 ton/acre
-   the mean almond yield anomaly should be approximately 182 ton/acre

### Load libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
```

### Load data

```{r}
climate <- read_table(here("data", "clim.txt")) %>% 
  clean_names()
```

### Almond Model

We linear regression equation that represents the relationship between the variables we are interested in. This equation is to calculate yield at a given point, i.e. a specific day, month, year. So, we will have to create a for loop that iterates this equation over all the data, then aggreagate to the min, max, and mean.

```{r}
# Variables:
# Y = Yield Anomaly 
# Tn = Min Temp
# Tx = Max Temp
# P = Precipitation

# Equation: Y = -0.015Tn - 0.0046Tn^2 - 0.07P + 0.0043P^2 + 0.28

almond_model <- function(climate) {
  
  # Filter climate data to only February
  climate <- climate %>% 
    filter(month == 2)
  
  # We need to calculate the variables in the model equation
  # I.e. Min temp and Precipation
  
  
}
```


The correct answer I think...

```{r}
# Group by year and month to get monthly averages
almond_month <- climate %>% 
  group_by(year, month) %>% 
  summarise(tmin_avg = mean(tmin_c),  
            tmax_avg = mean(tmax_c),
            precip_total = sum(precip)) %>% # you don't want average precip, you want total for the month!
  ungroup()
```

```{r}
# Filter to january and february because those are the months used in the paper
jan_data <- almond_month %>% 
  filter(month == 1) %>% 
  select(year, precip_total)

feb_data <- almond_month %>% 
  filter(month == 2) %>% 
  select(year, tmin_avg)

# Join just these two together by year
clean_almond <- left_join(feb_data, jan_data, by = "year")

# Fix the names
names(clean_almond) <- c("year", "tmin", "precip")
```

```{r}
almond_yield <- function(tmin, precip) {
  y <- (-0.015 * tmin - 0.0046 * tmin^2 - 0.07 * precip + 0.0043 * precip^2 + 0.28)
  return(y)
}

clean_almond$yield <- with(clean_almond, almond_yield(tmin, precip))
summary(clean_almond$yield)

```




This summary yields the exact numbers from the answer

```{r}
# OLD
almond_yield <- function(tmin, precip){
  y <- (-0.015*tmin - 0.0046*tmin**2 - 0.07*precip + 0.0043*precip**2 + 0.28)
  return(y)
}

mean_mint <- mean(climate$tmin_c)

almond_yield(mean_mint, mean(climate$precip))
```



```{r}
# yields <- tibble(
#   year = integer(),
#   month = integer(),
#   day = integer(),
#   yield = numeric()
# )

yields <- numeric(nrow(almond_month))


# Make year and month lists
years <- seq(range(climate$year)[1], range(climate$year)[2])
months <- seq(range(climate$month)[1], range(climate$month)[2])
days <- seq(range(climate$day)[1], range(climate$day)[2])


for (year in seq_along(years)){
  
  for (month in seq_along(months)){
    
    for (day in seq_along(days)) {
    
      row_data <- climate[day,]
    
      # Access values for min temp column
      tmin <- row_data$tmin_c
      
      # Access values for precipation column
      precip <- climate[ifelse(day == 1, 12, day - 1),]$precip
      
      # Use the equation to calculate almond yield for the specific observation
      y <- (-0.015*tmin - 0.0046*tmin**2 - 0.07*precip + 0.0043*precip**2 + 0.28)
      
      # store observation into a df for later use
      yields[i] <- y
      # yields$day <- day
      # yields$yield <- y
    }
    # yields$month <- month
  }
  # yields$year <- year
}

# Create a dataframe with the results
df <- tibble(yield = yields)

min(df$yield)
max(df$yield)
mean(df$yield)
```


```{r}
yields <- numeric(nrow(almond_month))

for (i in 1:nrow(almond_month)){
  
  # Access values for all columns 
  row_data <- almond_month[i,]
  
  # Access values for min temp column
  tmin <- row_data$mean_tmin
  
  # Access values for precipation column
  precip <- almond_month[i-1,]$mean_precip
  
  # Use the equation to calculate almond yield for the specific observation
  y <- (-0.015*tmin - 0.0046*tmin**2 - 0.07*precip + 0.0043*precip**2 + 0.28)
  
  # store observation into a df for later use
  yields[i] <- y
  
}

# Create a dataframe with the results
df <- tibble(yield = yields)

min(df$yield)

max(df$yield)

mean(df$yield)
```

