---
title: "Assignment 1: Creating a Model"
author: "Tom Gibbens-Matsuyama, Haylee Oyler"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

### Goal of this Assignment

Your goal is to implement a simple model of almond yield anomaly response to climate

- Inputs: daily times series of minimum, maximum daily temperatures and 
- Outputs: maximum, minimum and mean yield anomoly for a input time series

The Lobell et al. 2006 paper will be the source for the transfer function that you will use in your model; specifically
look at the equations in table 2.

Here are the expected outputs to check our work:

- the maximum almond yield anomaly should be approximately 1920 ton/acre
- the lowest almond yield anomaly should be approximately -0.027 ton/acre
- the mean almond yield anomaly should be approximately 182 ton/acre


### Load libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
```

### Load data
```{r}
climate <- read_table(here("data", "clim.txt")) %>% 
  clean_names()
```

