---
title: "Assignment 3: Profit Model"
author: "Tom Gibbens-Matsuyama, Haylee Oyler"
format: 
  html:
    embed-resources: true
editor: visual
execute:
  warning: false
  message: false
---

1.  Develop a profit model for your almond yield

-   Use your imagination to think about what would make sense
-   Make sure you include parameters
-   You could assume a baseline profit and then adjust according to the anomaly

2.  Apply the profit model to your almond yield model

There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function)

Almonds: - Corr: -0.71 - Slope: -1.04

3.  Perform a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

4.  Create 2 graphs

    -   one that shows yield anomaly for each year, accounting for uncertainty in the parameters
    -   one that show how yield anomaly in general varies with your parameters

5.  Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

Submit as a group: a Quatro document, your R files for almond yield and profit model on Canvas- Due Date Monday April 21

### Load libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(purrr)

source('R/compute_almond_yield.R')
source('R/compute_profit.R')
source('R/compute_NPV.R')
source('R/run_almond_profit_model.R')
```

### Load data

```{r}
climate <- read_table(here("data", "clim.txt")) %>% 
  clean_names()
```

### Apply the almond model

```{r}
run_almond_profit_model(data = climate, price=3000)
```

NEED TO ADD BASELINE YIELD TO MODEL From Ojas's email:

Just an update on the graphs, the intention was indeed to show almond profits and not yields. So it's up to you which one you would like to output. e.g. you can do it in the way we talked about in section:

-   One graph with almond yields through time with parameter uncertainty in yields
-   One graph with almond profits parameter uncertainty in profits (this is could be a combo of uncertainty in yield and profit parameters, or any one of the two, as long as it's at least two parameters). This could be overall for all years combined, or again through time.

You could also do it in another way:

-   One graph with almond yields in any one year or for all years together, with uncertainty
-   One graph with profits over time with uncertainty

Finally, you could also have both graphs with profits, overall uncertainty and through time. But it might be informative to see the yield results separately.

Feel free to pick and be creative if you feel like it, i'll accept all combinations, just please explain what you've done etc. as mentioned in the assignment.


## Sensitivity Analysis


### Profit Model
The two parameters we have chosen to vary for our profit model are the baseline price and the coefficient that represents the relationship between price and almond yield. We got the value of this coefficient from the original paper and we set the baseline price somewhat arbitrarily. We will refer to the coefficient as yield coefficient `byield` throughout the markdown. 

```{r}
# Set base values and deviation
deviation <- 0.15
base_price <- 4000
base_byield <- -1.04
n_samples <- 40

# Sample from uniform distributions
param_grid <- tibble(
  price = runif(n_samples, 
                base_price * (1 - deviation), base_price * (1 + deviation)),
  byield = runif(n_samples, 
                  range(base_byield * c(1 - deviation, 1 + deviation)))
)

```

```{r}
# Apply the profit function with the parameter grid above
profit_runs <- pmap(param_grid, function(price, byield) {
  run_almond_profit_model(climate, price = price, byield = byield) %>%
    mutate(price = price, byield = byield)
})

# Bind data together
profit_df_all <- bind_rows(profit_runs)

```

```{r}
# Graph results across time
profit_df_all %>% 
  # Drop 1995 because it skews the visualization
  filter(year != 1995) %>% 
  ggplot( aes(x = factor(year), 
                          y = netpresent_profit, group = year)) +
  geom_boxplot(fill = "lightgreen", color = "forestgreen") +
  labs(
    title = "Profit Distribution by Year with Price and Yield Coefficient Uncertainty",
    x = "Year", y = "Discounted Profit ($/acre)"
  ) +
  theme_minimal()

```
We decided to filter out the year 1995 because it was vastly skewing the results. That being said, many of the box plos are still hard to interpret because of the spread of the profit variable. This plot is still a little opaque, so we decided to isolate the effect of the two parameters more clearly in the following graphs.

```{r}
# Graph profit as a function of baseline price
p1 <- ggplot(profit_df_all, aes(price, netpresent_profit, col = byield)) +
  geom_point(cex = 2) +
  labs(y = "Net Present Profit ($)", 
       x = "Price ($/ton)",
       col = "Yield\nCoefficient",
       title = "Effect of Price and Almond Yield Coefficient\nUncertainty on Profit") +
  theme_minimal()
p1
```

```{r}
# Graph profit as a function of byield
p2 <- ggplot(profit_df_all, aes(byield, netpresent_profit, col = price)) +
  geom_point(cex = 2) +
  labs(y = "Net Present Profit ($)", 
       x = "Yield Coefficient",
       col = "Price",
       title = "Effect of Almond Yield Coefficient and Price\nUncertainty on Profit") +
  theme_minimal()
p2
```
Looking at the scatter plots of the relationship between profit and price and yield coefficient, we can see that price does not have a clear relationship with overall profit. This is somewhat surprising. However, `byield` does have a clear positive relationship with net profit, as seen in the graph above. This tells us that as the strength of the relationship between almond yield and price increases, the overall profit also increases. 

### Varying Almond Yield Parameters

For the almond yield model, we decided to vary the coefficient that represents the relationship between temperature and almond yield `btemp` as well as the coefficient that represents the relationship between precipitation and almond yield `brecip`. 

```{r}
# Set base values and deviation
deviation <- 0.15
base_btemp <- -0.015
base_bprecip <- -0.07
n_samples <- 40

# Using range here because the negative numbers were throwing NAs
param_grid <- tibble(
  btemp = runif(n_samples, 
                range(base_btemp * c(1 - deviation, 1 + deviation))),
  bprecip = runif(n_samples, 
                  range(base_bprecip * c(1 - deviation, 1 + deviation)))
)

```

```{r}
# Apply the parameter grid to the almond yield function
almond_runs <- pmap(param_grid, function(btemp, bprecip) {
  compute_almond_yield(climate, btemp = btemp, bprecip = bprecip) %>%
    mutate(btemp = btemp, bprecip = bprecip)
})

# Combine rows together
almond_df_all <- bind_rows(almond_runs)
```

```{r}
# Plot the relationship between yield and btemp
p1 <- ggplot(almond_df_all, aes(btemp, almond_yield, col = bprecip)) +
  geom_point(cex = 2) +
  labs(y = "Almond Yield (ton/acre)", 
       x = "Temperature Coefficient",
       col = "Precipitation\nCoefficient",
       title = "Effect of Temperature and Precipitation Coefficient Parameter\nUncertainty on Almond Yield Anomaly")
p1
```
```{r}
# Plot the relationship between yield and bprecip
p1 <- ggplot(almond_df_all, aes(bprecip, almond_yield, col = btemp)) +
  geom_point(cex = 2) +
  labs(y = "Almond Yield (ton/acre)", 
       x = "Precipitation Coefficient",
       col = "Temperature\nCoefficient",
       title = "Effect of Precipitation and Temperature Coefficient Parameter\nUncertainty on Almond Yield Anomaly")
p1
```
Looking at the scatter plots of the relationship between almond yield and the temperature and precipitation coefficients, we can see that yield does not have a clear relationship with temperature. However, `bprecip` does have a clear positive relationship with almond yield, as seen in the graph above. This tells us that as the strength of the relationship between almond yield and precipitation increases, the overall yield also increases. 

## Conclusion

Overall, our sensitivity analysis has shown that different parts of our model are selectively sensitive to different parameters. It is important to understand what drives the variability in your model's outcomes. In this example, we see that profit is driven primarily by `byield` over the baseline price, and almond yield is driven by `bprecip` over `btemp`. 

