---
title: "Assignment 3: Profit Model"
author: "Tom Gibbens-Matsuyama, Haylee Oyler"
format: html
editor: visual
---

1. Develop a profit model for your almond yield

- Use your imagination to think about what would make sense
- Make sure you include parameters
- You could assume a baseline profit and then adjust according to the anomaly

2. Apply the profit model to your almond yield model

There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function)

Almonds:
- Corr: -0.71
- Slope: -1.04

Almond func:
- The entire almond yield for that year

Profit func:
- Compute profit given yearly almond yield (pa)

profit_almond <- compute_profit_fromyield(
  almond_yield = clean_monthly$almond_yield, # input
  year = clean_monthly$year, # input
  price = 30, # idk (parameter?)
  discount = 0.04 # parameter
)

3. Perform  a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

4. Create 2 graphs
    - one that shows yield anomaly for each year, accounting for uncertainty in the parameters
    - one that show how yield anomaly in general varies with your parameters

5. Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

Submit as a group: a Quatro document, your R files for almond yield and profit model on Canvas- Due Date Monday April 21

### Load libraries

```{r}
library(tidyverse)
library(here)
library(janitor)

source('R/almond_yield.R')
```

### Load data

```{r}
climate <- read_table(here("data", "clim.txt")) %>% 
  clean_names()
```

### Apply the almond model

```{r}
almond_yield(data = climate)
```

